\chapter{Team charter checklist}
\label{appendix:team_charter_checklist}
\tfaq[1]{0mm}{What issues should be covered when drafting a team charter?}{Project~management}
Each person tends to have their own habits, work practices, and preferences.  These differences between people can cause challenges and tension within the team.  To avoid this, it is good practice to agree a set of rules of engagement or work practices in the form of a team charter\index{team~charter}.  To develop a team charter, the team must agree on a common set of practices and rules of engagement, as well as approaches to deal with deviations from the agreement, disagreements, and conflicts.  These are then written down in a document that can be referred to if concerns, disagreements, or conflicts arise.  By anticipating potential challenges and agreeing approaches to deal with these, such challenges can be dealt with more easily.

\begin{enumerate}
\item Communication
  \begin{itemize}
  \item Which tools should be used for day-to-day communication? (e.g. Slack, WhatsApp, Teams, email, etc.)
  \item What is the expected response time for messages?
  \item How will be handle urgent v non-urgent messages?
  \item We are required to organise regular team meetings (same time and place each week)
    \begin{itemize}
    \item How many time slots will we reserve per week? (e.g. one, \emph{two}, \emph{more})
    \item How will we meet? (e.g. \emph{in-person}, Teams, Zoom, etc.)
    \item Which time slots will be used every week?
    \item Who is responsible for convening/scheduling extra meetings when we need them?
    \end{itemize}
  \end{itemize}
\item Roles and responsibilities
  \begin{itemize}
  \item How will we assign roles? (e.g. project manager, team meeting chairperson, team meeting secretary/minute taker, client/stakeholder liaison, product owner, etc.)
  \item Will roles be fixed or rotated?
  \item Who is responsible for deadlines, progress tracking, and reporting? 
  \item (if relevant) Who is responsible for communication/engagement with the client?
  \item What are team member responsibilities in the days before the deadline to ensure that the deliverables are submitted on time?
  \end{itemize}
\item Collaboration tools
  \begin{itemize}
  \item Which GitHub service will we  use? (GitHub.com or GitHub.kcl.ac.uk)
  \item Where will we store and share documents? (e.g. GitHub Wiki, OneDrive, Google Drive, etc.)
  \item What project management tool will we use? (e.g. Trello, Jira, Notion, etc.)
  \end{itemize}
\item Decision making
  \begin{itemize}
  \item How will we make decisions? (e.g. majority vote, \emph{consensus}, role based)
  \item How will we resolve disagreements?
  \item Who has final say if there is a deadlock?
  \end{itemize}
\item Code development practices
  \begin{itemize}
  \item What coding standards and guides will we follow, and to what extent? (e.g. specify a subset of the code guidelines in the module handbook)
  \item How often should code be committed/pushed to version control? (e.g. at least once per day of code, push after every commit)
  \item What process do we use to review a branch before merging with the main?
  \item How will we ensure code quality? (e.g. pair programming, code review before merging, weekly/fortnightly code inspections)
  \item How will we track tasks? (e.g. Kanban board, meeting minutes, etc.)
  \item What is the expected level of automated testing?
  \end{itemize}
\item Accountability
  \begin{itemize}
  \item How will we track individual contributions? (e.g. GitHub, Team Feedback)
  \item How will we deal with missed deadlines? 
  \item How will we deal with work that fails to meet agreed standards?
  \item How will we deal with poor participation/engagement?
  \item Under what circumstances will the team reassign a task?
  \item What do we do with task assignment and incomplete tasks if a team member fails to attend a team meeting?
  \item If someone cannot contribute due to circumstances beyond their control, they should submit a \ac{mcf}.  How will you deal with such circumstances internally?
  \item How do we ensure that collaborative coding sessions are recorded correctly by the committer?
  \end{itemize}
\item Conflict resolution
  \begin{itemize}
  \item How will we raise concerns with the group?
  \item What steps will we take if there is a disagreement in the team?
  \item Under what circumstances should issues be escalated (e.g. to an advisor or the lecturer)?
  \end{itemize}
\item Availability and commitments
  \begin{itemize}
  \item What are the time constraints of each member?
  \item How many hours per week should each member expect to commit to the project?
  \item When should team members NOT be contacted? (e.g. weekdays 5pm-9am, weekends)
  \item How will we handle reading week and deadlines for other coursework?
  \end{itemize}
\item Team culture and values
  \begin{itemize}
  \item How will we ensure respect and inclusivity in discussion?
  \item How will we celebrate milestones and achievements?
  \item What behaviours are unacceptable? (e.g. ignoring messages, missing meetings without notice, etc.)
  \item How will the team respect and accommodate personal identities and needs?  What issues affect this particular team? (e.g. pronouns, culturally sensitive issues, disabilities, etc.)
  \end{itemize}
\item Review: When and how will the team accommodate the need for adjustments and changes to the charter?
\end{enumerate}