\begin{comment}
& Functionality & The team's web application contains one or more forms that enable a user to input data in the database.\\
\unit[30]{\%} & Functionality & The team has delivered a system that possesses a modest range of features (relative to the team's effective size), that contribute towards achieving the project's objectives  The software possesses some useful features.\\*
& Design & The software includes a Django model, view, and template.\\*
& Documentation & The software comes with the required \texttt{README.md} and \texttt{self-assessment.pdf} files that contains all required information.\\
\unit[40]{\%} & Functionality & The team delivered an application with an adequate breadth of features, given the team's effective size.  These contribute towards achieving project objectives.\\*
& Code & The team delivered software that is largely free from bugs.\\*

& Deployment & A version of the application is deployed and accessible through the access credentials specified in the assignment.  Where certain functionality is implemented but not usable through the deployed version (e.g. because of bugs that only the deployed version possesses, these have been identified clearly in the README.md file of the submission).\\
\unit[50]{\%} & Functionality & The team delivered software that enables the end-user achieve modest set of objectives, relative to the team's effective size.\\*
& Design & The teams has designed a sensible database structure and specified it through Django models.  The database accurately reflects the data required by the application.\\*
& Code & The team made a demonstrable effort to review and refactor code, ensuring good use of comments.  Each class, function, and method that is accessed called by other modules (i.e. from other files or classes), is documented with descriptive comments.  All "TODO" comments have been removed from the final version.  All commented out code has been removed.\\*
& Testing & The software comes with a test suite with good statement and branch coverage. At least 90\% of tests pass.\\*
& Deployment & All the application's features work in both the development and the deployed version of the software.   There are no features that have not been deployed.\\*
& Data seeding & The deployed version of the software is seeded with a substantial database, containing a sufficient volume records to perceive the application at scale.  The source code comes with a database seeder that generates a sufficient volume of data to experience the application at scale.\\*
& Management & The team's submission includes all required deliverables in full.\\
\unit[60]{\%} & Functionality & The software's \ac{UI} benefits from the use of a \ac{CSS} framework, such as Bootstrap\index{Bootstrap}.  The UI is consistent throughout.  There is evidence of coordination between members as screens have a similar look, feel, and behaviour across features.\\*
& Design & Large files by replacing the \texttt{views.py} and \texttt{models.py} files by directories containing small, well-labelled files with subsets of views and models.\\*
& Code & All names in the source code have a clear meaning, and make meaningful distinctions.\\*
& & The code is free from bugs.\\*
& Testing & The software comes with a test suite with excellent statement and branch coverage.  At least 99\% of tests pass.\\
\unit[70]{\%} & Functionality & The team delivered software that enables the end-user achieve a moderately ambitious set of objectives, relative to the team's effective size.\\*
& Functionality & The \ac{UI} is polished.  Examples include, but are not limited to: consistent use of user-friendly presentation of dates and times in UK format, the ability to retrieve records without having to know primary keys (unless that is explicitly required), large lists are paginated, lists contain links to key operations available to items, and related information is hyperlinked.\\*
& Design & View functions and methods of view classes are small and have limited responsibility.  Work is delegated to models, forms, helpers, and other components where appropriate.\\*
& & Refactoring and effective use of partials and template inheritance has prevented duplication of template code\\*
& Code & There are no large functions.  Function in the source code do one thing and one thing only. There are no functions with more than two levels of nesting.\\*
& & Code is mostly \ac{DRY}.\index{DRY~code}  Significant and obvious code repetition has been avoided.\\*
& Testing & The software comes with a test suite with impeccable statement and branch coverage.  All tests pass.\\
\unit[80]{\%} & Functionality & All screens involving the presentation of lists have been developed in-depth and incorporate some means to find information more easily.  These may include means to search records, or sort lists based on a field.\\*
& Design & The software makes good use of class-based and class-based generic views\index{Django!class-based views} to avoid unnecessary reimplementation of \ac{CRUD} operations.\\*
& Code & All code is \ac{DRY}.  There is no code repetition.\\*
& Testing & Inspection of the test code indicates excellent coverage of all meaningful input and output partitions.\\
\unit[90]{\%} & Functionality & All screens involving the presentation of lists have been developed in-depth and incorporate comprehensive search, and ordering features. \\*
& Design & The software makes extensive use of class-based and class-based generic views to avoid any repetition of control logic in views.\\*
& Code & Functions and methods are very short and focussed. No functions or methods have more than one level of nesting.\\*
& Testing & The submission includes a suite of functional tests (e.g. using Selenium)\index{Selenium} covering a good range of typical use cases.\\
\unit[100]{\%} & Functionality & The team delivered software that enables the end-user achieve a very ambitious set of objectives.  The scope that has been achieved is exceptional, relative to the team's effective size.\\*
& Design \& code & Design quality and code cleanliness are exemplary.\\*
& Testing & The submission includes a comprehensive set of functional tests (e.g. using Selenium) covering all use cases.\\*
& Management & The team's time management has been excellent.  The final five days of the project were free from significant development activity so that the team could focus on refactoring the previous week's work and quality control of the deliverables.
\end{comment}